Reanalysis Step 5: Outcome Analysis with Three-Way Interaction

### 2025-09-01 reanalysis-5-outcome_analysis-interaction.qmd
clean up ix

if (file.exists(models_cache_file) && !force_refit_models) {
  # Load cached models
  cached_models <- readRDS(models_cache_file)
  models <- cached_models$models
  model_summaries <- cached_models$model_summaries
  model_progress <- cached_models$model_progress
  models_loaded_from_cache <- TRUE
  cache_location <- models_cache_file
  message("Models loaded from cache: ", cache_location)
} else {
  # Fit new models
  models_loaded_from_cache <- FALSE
  models <- list()
  model_summaries <- list()

@@

  saveRDS(cached_models, models_cache_file)
  cache_location <- models_cache_file
  message("Models fitted and saved to cache: ", cache_location)
}

@@

@@

@@

  # Check column names in first imputation for debugging
  if (length(all_contrasts) > 0) {
    message("Contrast columns: ", paste(names(all_contrasts[[1]]), collapse = ", "))
  }
  if (length(all_risk_diffs) > 0) {
    message("Risk diff columns: ", paste(names(all_risk_diffs[[1]]), collapse = ", "))
  }

  # Pool each subgroup using Rubin's rules
  for (j in 1:n_subgroups) {
    # Pool odds ratio contrasts (log scale)
    # Extract the j-th estimate from each imputation
    log_odds_estimates <- sapply(all_contrasts, function(x) {
      x$estimate[j]
    })

@@

if (file.exists(emmeans_cache_file) && !force_recalc_emmeans) {
  # Load cached results
  pooled_results <- readRDS(emmeans_cache_file)
  emmeans_loaded_from_cache <- TRUE
  emmeans_cache_location <- emmeans_cache_file
  message("Emmeans results loaded from cache: ", emmeans_cache_location)
} else {
  # Calculate new emmeans results
  emmeans_loaded_from_cache <- FALSE

@@

  # Save to cache
  saveRDS(pooled_results, emmeans_cache_file)
  emmeans_cache_location <- emmeans_cache_file
  message("Emmeans results calculated and saved to cache: ", emmeans_cache_location)
}

@@

# Display the risk differences prominently
message("\n=== RISK DIFFERENCES (Probability Scale) ===")
print(rd_table)
```

### Understanding the Different Effect Measures

::: {.callout-note}
## Why are the confidence intervals so different between measures?

The three effect measures presented above capture different aspects of the lapsing effect:

1. **Odds Ratios (OR)**: 
   - Multiplicative scale comparing odds of outcome
   - Wider CIs because they incorporate uncertainty from multiple interaction terms
   - Non-collapsible measure that depends on covariate distribution

2. **Risk Differences (RD)**:
   - Additive scale showing absolute change in probability
   - Narrower CIs because probabilities are bounded [0,1]
   - More intuitive interpretation for clinical decision-making

3. **Average Marginal Effects (AME)** (shown later):
   - Average of individual-level effects across actual patients
   - Smallest CIs because averaging reduces variance
   - Accounts for actual covariate distribution in each subgroup

The OR confidence intervals are wider because:
- They must account for uncertainty in the main effect plus all relevant interaction terms
- The log scale amplifies uncertainty when exponentiating
- Small subgroups have fewer effective degrees of freedom
:::

@@

::: {.callout-note collapse="true"}
## Click to expand manual calculations

This section shows the manual calculation of subgroup effects for validation purposes.

@@

# Check if we have valid results
n_valid <- sum(!is.na(manual_effects$Log_Odds))
if (n_valid == 0) {
  message("Warning: No valid manual effect calculations")
} else if (n_valid < nrow(manual_effects)) {
  message("Warning: Some manual effect calculations contain NA values (", 
          nrow(manual_effects) - n_valid, " out of ", nrow(manual_effects), ")")
}

@@

:::

@@

message("Results saved to: ", 
        file.path(reanalysis_data_dir, "outcome_analysis_results.rds"))
```

>>>>

if (file.exists(models_cache_file) && !force_refit_models) {
  # Load cached models
  cached_models <- readRDS(models_cache_file)
  models <- cached_models$models
  model_summaries <- cached_models$model_summaries
  model_progress <- cached_models$model_progress
  models_loaded_from_cache <- TRUE
  cache_location <- models_cache_file
  log_message("Models loaded from cache: ", cache_location)
} else {
  # Fit new models
  models_loaded_from_cache <- FALSE
  models <- list()
  model_summaries <- list()

@@

  saveRDS(cached_models, models_cache_file)
  cache_location <- models_cache_file
  log_message("Models fitted and saved to cache: ", cache_location)
}

@@

  # Initialize progress tracking
  progress <- init_progress(n_imputations, "Calculating contrasts")

  for (i in 1:n_imputations) {
    # Update progress
    progress <- update_progress(progress)

@@

  # Finalize progress
  finalize_progress(progress)

@@

  # Check column names in first imputation for debugging
  if (length(all_contrasts) > 0) {
    log_message("Contrast columns: ", paste(names(all_contrasts[[1]]), collapse = ", "))
  }
  if (length(all_risk_diffs) > 0) {
    log_message("Risk diff columns: ", paste(names(all_risk_diffs[[1]]), collapse = ", "))
  }

  # Pool each subgroup using Rubin's rules
  log_progress(0, n_subgroups, "Pooling subgroup")

  for (j in 1:n_subgroups) {
    log_progress(j, n_subgroups, "Pooling subgroup")
    # Pool odds ratio contrasts (log scale)
    # Extract the j-th estimate from each imputation
    log_odds_estimates <- sapply(all_contrasts, function(x) {
      x$estimate[j]
    })

@@

if (file.exists(emmeans_cache_file) && !force_recalc_emmeans) {
  # Load cached results
  pooled_results <- readRDS(emmeans_cache_file)
  emmeans_loaded_from_cache <- TRUE
  emmeans_cache_location <- emmeans_cache_file
  log_message("Emmeans results loaded from cache: ", emmeans_cache_location)
} else {
  # Calculate new emmeans results
  emmeans_loaded_from_cache <- FALSE

@@

  # Save to cache
  saveRDS(pooled_results, emmeans_cache_file)
  emmeans_cache_location <- emmeans_cache_file
  log_message("Emmeans results calculated and saved to cache: ", emmeans_cache_location)
}

@@

# Display the risk differences prominently
log_message("Risk differences (probability scale) calculated")
```

@@

@@

# Check if we have valid results
n_valid <- sum(!is.na(manual_effects$Log_Odds))
if (n_valid == 0) {
  log_message("Warning: No valid manual effect calculations")
} else if (n_valid < nrow(manual_effects)) {
  log_message("Warning: Some manual effect calculations contain NA values (", 
          nrow(manual_effects) - n_valid, " out of ", nrow(manual_effects), ")")
}

@@

@@

log_message("Results saved to: ", 
        file.path(reanalysis_data_dir, "outcome_analysis_results.rds"))
```
